---
title: "WineCase"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Read in Data
```{r}
rm(list = ls())
library(dplyr)
library(tidyr)
library(data.table)
library(ggplot2)
library(grf)
setwd("~/Desktop/UR/Spring A/GBA424 Analytics Design and Applications/Case4 Wine Retailer Case 1904")
d = read.csv("test_data_1904.csv", stringsAsFactors = TRUE)
```

## Summary Stats for Treatment and Control
```{r}
dt = data.table(d)
dagg = dt[,.(open = mean(open), click=mean(click), purch = mean(purch),seOpen = sd(open)/sqrt(.N), seClick=sd(click)/sqrt(.N), sePurch = sd(purch)/sqrt(.N),.N),by = .(group)]
dagg
```

## Plot results of Purchase given the treatment or control
```{r}
dodge = position_dodge(width=1); ##to form constant dimensions
ggplot(aes(x=group,y=purch,ymax=purch+sePurch,ymin=purch-sePurch),data=dagg)+
  geom_bar(position=dodge,stat="identity",col=2:3,fill=2:3) + 
  geom_errorbar(position=dodge)
  labs(x="Group",y="Purchases")
  
##Even when taking standard errors into account, the purchase of each group, in dollar amounts
##does not overlap. Could this signify that the email has an impact on purchase amount?

```

```{r}
summary(lm(purch~group,data=d)) 
## We observe that the baseline purchase amount is $12.77
## If given the treatment of email, expect purchase to increase by $1.35, from the baseline of no email

##Will not add covariates as the treatment effect appears to be large and significant on its own.
##Worst case scenario: $1.02 increase in purchase
##Best case scenario: $1.66 increase in purchases
```

## Histogram of Effect given Purchase Recency (60 Days as recent purchase)
```{r}
hist(d$last_purch, 
     xlab="Days Since Last Purchase", ylab="Customers", 
     main="Histogram of Days Since Last Purchase")
d$recentPurch = (d$last_purch < 60)
dt = data.table(d)

## Will use 60 Days as a recent customer. May consider 65 to keep both recent and non recent group equal in 
## sample size
```

## Responses to treatment given whether or not they are recent buyers
```{r}
dagg = dt[,.(open = mean(open), click=mean(click), purch = mean(purch),seOpen = sd(open)/sqrt(.N), seClick=sd(click)/sqrt(.N), sePurch = sd(purch)/sqrt(.N),.N),by = .(group,recentPurch)]
dagg

##Effect of treatment given type of customer
# Recent Buyer: Increase of $1.74 when receive email
# Old Buyer: Increase of $0.92 when receive email
```

```{r}
dodge = position_dodge(width=1); ##to form constant dimensions
ggplot(aes(fill=group,y=purch,x=recentPurch,ymax=purch+sePurch,ymin=purch-sePurch),data=dagg)+
  geom_bar(position=dodge,stat="identity") + 
  geom_errorbar(position=dodge)
  labs(x="Group",y="Purchases")
  
## Still no overlap of effects, even when taking standard errors into account

```

## Slicing on different type of website visits
```{r}
dagg = dt[,.(open = mean(open), click=mean(click), purch = mean(purch),seOpen = sd(open)/sqrt(.N), seClick=sd(click)/sqrt(.N), sePurch = sd(purch)/sqrt(.N),.N),by = .(group, visits >5)]
dagg

## Visit is True : 1.17
## Visit is False: 1.50

##Intuition:
# Likely that those who visit t
```

```{r}
dodge = position_dodge(width=1); ##to form constant dimensions
ggplot(aes(fill=group,y=purch,x=visits,ymax=purch+sePurch,ymin=purch-sePurch),data=dagg)+
  geom_bar(position=dodge,stat="identity") + 
  geom_errorbar(position=dodge)
  labs(x="Group",y="Purchases")
  
## Still no overlap between email and no email, difference is a lot tighter for those who visit more recently
```

## Slicing on wine preference: Chardonay
```{r}
dagg = dt[,.(open = mean(open), click=mean(click), purch = mean(purch),seOpen = sd(open)/sqrt(.N), seClick=sd(click)/sqrt(.N), sePurch = sd(purch)/sqrt(.N),.N),by = .(group, chard >5)]
dagg

## No Chardonay: 1.16
## Buys Chardonay:1.686856

##Intuition:
# A majority of the people do not buy chardonay, and a lot less buy it. This is  then reflected in the
# standard errors. Even when taking the larger standard errors into account we observe that the groups are still
# different in treatment effect
# email has a bigger impact on individuals that buy chardonay...

```
```{r}
dodge = position_dodge(width=1); ##to form constant dimensions
ggplot(aes(fill=group,y=purch,x=chard,ymax=purch+sePurch,ymin=purch-sePurch),data=dagg)+
  geom_bar(position=dodge,stat="identity") + 
  geom_errorbar(position=dodge)
  labs(x="Group",y="Purchases")
  
  
```

## Slicing on wine preference: sauvignon blanc
```{r}
dagg = dt[,.(open = mean(open), click=mean(click), purch = mean(purch),seOpen = sd(open)/sqrt(.N), seClick=sd(click)/sqrt(.N), sePurch = sd(purch)/sqrt(.N),.N),by = .(group, sav_blanc >0)]
dagg

##Treatment effect... email
## No Sauvignon Blanc : $.75
## Buys Sauvignon Blanc: $2.80

##Intuition:
# The email treatment appears to have a greater impact on purchase for those that purchase sauvignon blanc
# Could this be as a result of the price, popularity, promotions for sauv blanc...
# More likely to try a product that is less expensive... 
# Standard errors are worrysome but expected due to sample size

```

```{r}
dodge = position_dodge(width=1); ##to form constant dimensions
ggplot(aes(fill=group,y=purch,x=sav_blanc,ymax=purch+sePurch,ymin=purch-sePurch),data=dagg)+
  geom_bar(position=dodge,stat="identity") + 
  geom_errorbar(position=dodge)
  labs(x="Group",y="Purchases")
  
```

## Slicing on wine preference: syrah 
```{r}
dagg = dt[,.(open = mean(open), click=mean(click), purch = mean(purch),seOpen = sd(open)/sqrt(.N), seClick=sd(click)/sqrt(.N), sePurch = sd(purch)/sqrt(.N),.N),by = .(group, syrah >0)]
dagg

#email is only effective for those that do not buy syrah
# Does not buy syrah : $1.46871

```

```{r}
dodge = position_dodge(width=1); ##to form constant dimensions
ggplot(aes(fill=group,y=purch,x=syrah,ymax=purch+sePurch,ymin=purch-sePurch),data=dagg)+
  geom_bar(position=dodge,stat="identity") + 
  geom_errorbar(position=dodge)
  labs(x="Group",y="Purchases")
  
```

## Slicing on wine preference: syrah 
```{r}
dagg = dt[,.(open = mean(open), click=mean(click), purch = mean(purch),seOpen = sd(open)/sqrt(.N), seClick=sd(click)/sqrt(.N), sePurch = sd(purch)/sqrt(.N),.N),by = .(group, cab >0)]
dagg

#Treatment effect on cabernet buyers
# Does not buy Cabernet : $1.08651
# Buys Carbenet: $1.95

##email may be more effective for cab lovers


```

```{r}
dodge = position_dodge(width=1); ##to form constant dimensions
ggplot(aes(fill=group,y=purch,x=cab,ymax=purch+sePurch,ymin=purch-sePurch),data=dagg)+
  geom_bar(position=dodge,stat="identity") + 
  geom_errorbar(position=dodge)
  labs(x="Group",y="Purchases")
  
```

## Repeated significance testing
## Measuring causal effects with regression: Conditional causal effects
## 1. last_purch
```{r}
summary(lm(purch~0+recentPurch + group:recentPurch,data=d))
```
## 2. visits
```{r}
d$highvisits = (d$visits >5)
summary(lm(purch~0+highvisits+group:(highvisits),data=d))
```
## 3. past_purch
```{r}
d$highpast_purch = (d$past_purch >60)
summary(lm(purch~0+highpast_purch+group:(highpast_purch),data=d))
```
## 4. chard
```{r}
d$boughtchard = (d$chard >0)
summary(lm(purch~0+boughtchard+group:(boughtchard),data=d))
```
## 5. sav_blanc
```{r}
d$boughtsav = (d$sav_blanc >0)
summary(lm(purch~0+boughtsav+group:(boughtsav),data=d))
```
## 6. syrah
```{r}
d$boughtsyrah = (d$syrah >0)
summary(lm(purch~0+boughtsyrah+group:(boughtsyrah),data=d))
```
## 7. cab
```{r}
d$boughtcab = (d$cab >0)
summary(lm(purch~0+boughtcab+group:(boughtcab),data=d))
```

## Incrementality: Lift aka Uplift model for purchase amount 
```{r}
d$email = (d$group != "ctrl")
m <- lm(purch ~ email*(last_purch < 60) + email*(past_purch > 60) + email*(visits > 5) +
                 email*(chard > 0) + email*(sav_blanc>0)  + email*(cab>0), 
         data=d)
summary(m)$coef
```


## Scoring customers with an uplift model
If you have someone who wasn't in the test, but you know their baseline variables, you can use an uplift model to predict likely treatment effect. 
```{r}
new_cust <- data.frame(chard=rep(38.12, 2), sav_blanc=rep(0, 2), 
                       syrah=rep(0, 2), cab=rep(0, 2),  
                       past_purch=rep(38.12,2), last_purch=rep(19,2), 
                       visits=rep(3,2))
(pred <- predict(m, cbind(email=c(TRUE, FALSE), new_cust)))
(lift <- pred[1] - pred[2])
```


## Causal forests
## training model
```{r}
set.seed(1)
cf_size <- 60000; #nrow(d) 
cf_set = sample(nrow(d),cf_size)
treat <- d$email[cf_set]
response <- d$purch[cf_set]
baseline <- d[cf_set, c("last_purch", 'past_purch',"visits", "chard", "sav_blanc",  "cab")]
tmp=proc.time()[3]
cf <- causal_forest(baseline, response, treat)
tmp = proc.time()[3]-tmp
print(cf)
```

## Overall average treatment effect
```{r}
average_treatment_effect(cf, method="AIPW")
```

## Predicted uplift 
Just like any uplift model, we can use the model to predict the email effect for new customers. 
```{r}
new_cust <- data.frame(chard=38.12, sav_blanc=0, syrah=0, cab=0, last_purch=19, visits=3)
predict(cf, new_cust, estimate.variance = TRUE)
```
## Predicted uplift for all customers in training set
```{r}
hist(predict(cf)$predictions, 
     main="Histogram of Purchase Lift", 
     xlab="Purchase Lift for Email", ylab="Customers")
```

```{r}
baseline$pred = predict(cf, baseline)$predictions
baseline$score <- baseline$pred*0.3 - 0.1
## the costomers with score > 0.1 will be our targeted customers
write.csv(baseline,'training.csv')
percent <- nrow(baseline[baseline$score > 0,])/60000
percent
```


## Uplift versus past purchase amount
```{r}
trans_gray <- rgb(0.1, 0.1, 0.1, alpha=0.1)
plot(d$past_purch[1:cf_size], predict(cf)$predictions, 
     cex=0.5, col=trans_gray,
     xlab="Past Purchase Amount ($)", ylab="Predicted Treatment Effect ($)")
```

## Uplift versus days since last purchase
```{r}
trans_gray <- rgb(0.1, 0.1, 0.1, alpha=0.1)
plot(d$last_purch[1:cf_size], predict(cf)$predictions, 
     cex=0.5, col=trans_gray,
     xlab="Days Since Last Purchase", ylab="Predicted Treatment Effect ($)")
```
## Test model using test set
```{r}
test = d[-cf_set, c("last_purch", 'past_purch',"visits", "chard", "sav_blanc",  "cab")]
test$pred = predict(cf, test)$predictions
```

## Predict lift using entire set
```{r}
entire = d[,c("last_purch", 'past_purch',"visits", "chard", "sav_blanc",  "cab")]
entire$pred = predict(cf,entire)$predictions
```

## Summary of target customers
```{r}
entire$target=entire$pred>(.10/.3)
targetcus = entire[entire$target==TRUE,]
summary(targetcus)
```

## Evaluate segmentation amongst those chosen for target email
```{r}
entireTRUE = entire[entire$target==TRUE,]
lastPurchTRUE = nrow(entireTRUE[entireTRUE$last_purch<60,])/nrow(entireTRUE)
pastPurchTRUE = nrow(entireTRUE[entireTRUE$past_purch>60,])/nrow(entireTRUE)
visitsTRUE = nrow(entireTRUE[entireTRUE$visits>5,])/nrow(entireTRUE) 
chardTRUE = nrow(entireTRUE[entireTRUE$chard>0,])/nrow(entireTRUE)
savBlancTRUE = nrow(entireTRUE[entireTRUE$sav_blanc>0,])/nrow(entireTRUE)
cabTRUE = nrow(entireTRUE[entireTRUE$cab>0,])/nrow(entireTRUE)

segTRUE = c("last purch"=lastPurchTRUE,"past purch"=pastPurchTRUE,
            "visits"= visitsTRUE, "chard"=chardTRUE,"sav_Blanch" = savBlancTRUE,
            "cab" = cabTRUE)
```

```{r}
entireFALSE = entire[entire$target==FALSE,]
lastPurchFALSE = nrow(entireFALSE[entireFALSE$last_purch<60,])/nrow(entireFALSE)
pastPurchFALSE = nrow(entireFALSE[entireFALSE$past_purch>60,])/nrow(entireFALSE)
visitsFALSE = nrow(entireFALSE[entireFALSE$visits>5,])/nrow(entireFALSE) 
chardFALSE = nrow(entireFALSE[entireFALSE$chard>0,])/nrow(entireFALSE)
savBlancFALSE = nrow(entireFALSE[entireFALSE$sav_blanc>0,])/nrow(entireFALSE)
cabFALSE = nrow(entireFALSE[entireFALSE$cab>0,])/nrow(entireFALSE)

segFALSE = c("last purch"=lastPurchFALSE,"past purch"=pastPurchFALSE,
            "visits"= visitsFALSE, "chard"=chardFALSE,"sav_Blanch" = savBlancFALSE,
            "cab" = cabFALSE)
```

```{r}
segTRUE
segFALSE
seg = as.data.frame(rbind(segFALSE,segTRUE))
```

## Predict target email for new customer
```{r}
##newChard, newSav, newSyrah, newCab: dollar amount spent per type of wine; Input 0 if not purchased
##newLastPurch: the number of days since last purchase.
##newPastPurch: dollar amount of previous purchase
##newVisits: number of visits to website by individual

##################################################
##### Input Values to evaluate new customer ######
##################################################
newChard = 29
newSav = 54
newCab = 98
newLastPurch = 20
newPastPurch = 12
newVisits = 6
##################################################
##### Input Values to evaluate new customer ######
##################################################

new_cust <- data.frame(chard=newChard, sav_blanc=newSav, cab=newCab, 
                       last_purch= newLastPurch, past_purch = newPastPurch ,visits=newVisits)
value = (predict(cf, new_cust, estimate.variance = TRUE))
if(value[1] > (.1/.3)){
  print("Send Target Email")
}else{
  print("Do not send Target Email")
}

```

